DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-08 Draft-Ped-Metrics-Example.Rmd
change scoring to allow for sidewalks when shoulder is standard

- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards)
- 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
- 1 if the sidewalk or shoulder is present and is narrower than AASHTO
- 0 if none present or no data

@@

```{r}
score_interpretation <- tribble(
  ~score_at_least, ~interpretation,
  10, "Good",
  5, "Adequate",
  3, "Minimum",
  0, "Poor",
)

@@

getSideScore <- function(standard_sidewalk, standard_shoulder,
                         sidewalk, shoulder) {
  ifelse(is.na(standard_sidewalk), # dealing with shoulder
         case_when(
         shoulder >= standard_shoulder ~ 5,
         shoulder < standard_shoulder & shoulder > 4 ~ 3,
         shoulder < standard_shoulder & shoulder > 0 ~ 1,
         TRUE ~ 0
         ),
         case_when(
         sidewalk >= standard_sidewalk ~ 5,
         sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
         sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
         TRUE ~ 0
         )
  )
}

@@

Some of the segments with score of 0 are because they have shoulders when they should have sidewalks. Here's a few example segments. If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate.

@@

roads_0 <- comb_score2 %>%
  filter(score == 0) %>%
  select(setting, road_class, FULL_NAME_R, sidewalk_right:score)

knitr::kable(head(roads_0)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road. A few details for roads where at least one segment has a 10 are shown here. 
```{r}
roads_10 <- comb_score2 %>%
  filter(score == 10) %>%
  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
  summarise(segments = nice_comma(n()), 
            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop")

@@

@@

```{r loop_libraries,include=FALSE,echo=FALSE}
the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)

@@

```{r loop_transit}
the_transit <- unique(st_set_geometry(transit_cp2, NULL)$id)
loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_transit, Score = NA))

expanded_child <- lapply(
    the_transit,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )
    })

@@

```{r}
score_breakdowns %>%
  rowwise() %>%
  mutate(Interpretation = getScoreInterpretation(Score))

knitr::kable(score_breakdowns) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

In addition, many roads have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**

>>>>

- 5 if it has a sidewalk or shoulder of adequate size (based on road classification & standards) or a sidewalk is present where a shoulder is the standard
- 3 if the sidewalk or shoulder is present and is minimum AASHTO requirement (like a 4 ft shoulder)
- 1 if the sidewalk or shoulder is present and is narrower than AASHTO (**NOTE to subcommittee - do we think we should use a 1 if there is a shoulder when there should be a sidewalk?**)
- 0 if none present or no data

@@

```{r}
score_interpretation <- tribble(
  ~score_at_least, ~interpretation,
  10, "Good",
  5, "Adequate",
  3, "Minimum",
  1, "Poor",
  0, "Nothing"
)

@@

getSideScore <- function(standard_sidewalk, standard_shoulder,
                         sidewalk, shoulder) {
  ifelse(is.na(standard_sidewalk), # dealing with shoulder
         case_when(
           !is.na(sidewalk) ~ 5, # there is a sidewalk present when a shoulder is the standard
           shoulder >= standard_shoulder ~ 5,
           shoulder < standard_shoulder & shoulder > 4 ~ 3,
           shoulder < standard_shoulder & shoulder > 0 ~ 1,
           TRUE ~ 0
         ),
         case_when(
           sidewalk >= standard_sidewalk ~ 5,
           sidewalk < standard_sidewalk & sidewalk > 4 ~ 3,
           sidewalk < standard_sidewalk & sidewalk > 0 ~ 1,
           TRUE ~ 0
         )
  )
}

@@

Some of the segments with score of 0 are because they have shoulders when they should have sidewalks. If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate. Here's a few example segments.

@@

roads_0 <- comb_score2 %>%
  filter(score == 0) %>%
  select(setting, road_class, FULL_NAME_R, sidewalk_right:standard_sidewalk)

knitr::kable(head(roads_0) %>%
               rename(Setting = setting,
                      Class = road_class,
                      Name = FULL_NAME_R,
                      `sidewalk R` = sidewalk_right,
                      `sidewalk L` = sidewalk_left,
                      `shoulder R` = shoulder_right,
                      `shoulder L` = shoulder_left,
                      `standard sidewalk` = standard_sidewalk)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road. A few details for roads where at least one segment has a 10 are shown here. 
```{r}
roads_10 <- comb_score2 %>%
  filter(score == 10) %>%
  group_by(setting, road_class, FULL_NAME_R) %>%
  summarise(segments = nice_comma(n(),0), 
            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop") %>%
  rename(Setting = setting,
         Class = road_class,
         Name = FULL_NAME_R,
         `Segments w 10` = segments,
         `Segment Length` = road_len,
         `Ped Fac Length` = dist)

@@

### Commissioner Districts

```{r plot_commdists}
commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
commdists <- st_as_sf(commdists) %>% 
  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
                ifelse(DISTRICT == "2", "2 South", "3 Central")))

kitsap +
  geom_sf(data = commdists, aes(fill = NAME), alpha = 0.5) +
  theme_void() +
  labs(title = 'Kitsap County Commissioner Districts',
       caption = "Data source: Kitsap GIS - data in development",
       color = "DIMS")

```{r loop_commdists,include=FALSE,echo=FALSE}
the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
loop_df <- st_as_sf(commdists) %>% mutate(tomatch = NAME)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_commdists, Score = NA))

expanded_child <- lapply(
    the_commdists,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )

Kitsap County has `r length(the_commdists)` Commissioner Districts.

@@

```{r loop_libraries,include=FALSE,echo=FALSE, cache = TRUE}
the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)

@@

```{r loop_transit,include=FALSE,echo=FALSE, cache = TRUE}
the_transits <- unique(st_set_geometry(transit_cp2, NULL)$id)
loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_transits, Score = NA))

expanded_child <- lapply(
    the_transits,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )
    })

@@

score_breakdowns <- score_breakdowns %>%
  rowwise() %>%
  mutate(Interpretation = getScoreInterpretation(Score))
write_csv(score_breakdowns, paste0(this_dir,"score_breakdowns.csv"))

getScoreBreakdownForArea <- function(the_area) {
  knitr::kable(score_breakdowns %>% 
                 filter(Area %in% eval(as.symbol(the_area))) %>%
                 mutate(Score = nice_comma(Score,2) 
                                           ),
             caption = paste("Scores for",
                             str_to_title(sub("the_"," ",the_area))
                             )
             ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
}

```{r, echo=FALSE, results='asis'}
score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
  knitr::knit_child(text = c(
    '## `r gsub("_"," ",x)`',
    '',
    '```{r}',
    'getScoreBreakdownForArea(x)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')

@@

In addition, many road segments have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**

### Notes to the special committee drafting this
We have a few to dos on this before we move forward.

1. Be sure that our criteria make sense to us
2. Consider how the scores should be interpreted - does the interpretation fit?
3. Consider if we are missing something critical
